\thetitle

### 2021-03-04 writing/rational-natural.tex
dangling lines
uniform editorial voice, mistakes, typos, size of text in figures, captions

The Natural semantics assigns blame to exactly one boundary.  A blame assignment
has the following specific meaning: the typed module may make incorrect type
assumptions about the untyped module in its interface, or the correct
interface exposes a bug in the untyped module (or its dependencies). Our setup rules out the first
alternative (but see section~\ref{sec:conclusion}), and therefore the rational
programmer extends the trail to a scenario that swaps out the untyped
module
for its typed counterpart.

Formally,
\begin{quote}
\it A \emph{Natural blame trail} is a sequence of scenarios $\conf_0,...\conf_n$ of
a program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i =\{\blame{\system, \conf_i}\}$, where
$\blame{\system, \conf}$ denotes the module that $\conf$ blames.
\end{quote}

@@

To account for this situation, a new mode of the Natural rational
programmer follows a migration process based on exceptions:
\begin{quote}
\it A {\em Natural exception trail\/} is a sequence of scenarios $\conf_0,...\conf_n$ of a
program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i = \{\exception{\system, \conf_i}\}$
where $\exception{\system, \conf}$ denotes the location of the exception of $\conf$.
\end{quote}

>>>>

The Natural semantics assigns blame to exactly one boundary.  A blame assignment
has the following specific meaning: the typed module may make incorrect type
assumptions about the untyped module in its interface, or the correct
interface exposes a bug in the untyped module (or its dependencies). Our setup rules out the first
alternative (but see sec.~\ref{sec:conclusion}), and therefore the rational
programmer extends the trail to a scenario that swaps out the untyped
module
for its typed counterpart.

% Formally, %% <-- MF: this is called a ``dangling def.'' and is considered bad style

Here is a rigorous definition. 
\begin{quote}
\it A \emph{Natural blame trail} is a sequence of scenarios $\conf_0,...\conf_n$ of
a program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i =\{\blame{\system, \conf_i}\}$, where
$\blame{\system, \conf}$ denotes the module that $\conf$ blames.
\end{quote}

@@

To account for this situation, a new mode of the Natural rational
programmer follows a migration process based on exceptions.
\begin{quote}
\it A {\em Natural exception trail\/} is a sequence of scenarios $\conf_0,...\conf_n$ of a
program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i = \{\exception{\system, \conf_i}\}$
where $\exception{\system, \conf}$ denotes the location of the exception of $\conf$.
\end{quote}

### 2021-03-04 writing/rational-transient.tex
dangling lines
uniform editorial voice, mistakes, typos, size of text in figures, captions

The Transient semantics assigns blame to a set of modules. The blame
assignment says that the value witnessing the impedance mismatch may have
crossed the boundaries between elements in the set, and each crossing
checked the value's type in a shallow manner.

This ambiguity in Transient blame raises the question of how the rational
programmer should react when the language produces a blame set. Our answer
is that the rational programmer has at least two reasonable options. The first
one is to select the module that is added to the blame set first and
assign types to only that one---after all, if fully checked, the types of this
first module should be able to detect an impedance mismatch earlier in the
evaluation of a program than the later ones. The second option is to select the
module that is added to the blame set last, effectively interpreting the
blame set as a boundary-aware stack.

@@

\begin{quote}
\it A \emph{Transient-first blame trail} is a sequence of scenarios
$\conf_0,...\conf_n$ of $\system$ where for all $0 \leq i \leq n - 1$,
$\conf_i \subset \conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i =
\first{\mblame{\system, \conf_i}}$ where $\first{\mblame{\system, \conf}}$ is the
first module Transient adds to the blame set for $\conf$.
\end{quote}

\begin{quote}
\it A \emph{Transient-last blame trail} is a sequence of scenarios
$\conf_0,...\conf_n$ of $\system$ where for all $0 \leq i \leq n - 1$,
$\conf_i \subset \conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i =
  \last{\mblame{\system, \conf_i}}$ where $\last{\mblame{\system, \conf}}$ is the
last module Transient adds to the blame set for $\conf$.
\end{quote}

The definition of \emph{Transient exception trails} is analogous to that for
Natural. It is used to specify when Transient-first and Transient-last blame are
more useful than Transient exceptions for debugging a scenario.

>>>>

The Transient semantics assigns blame to a sequence of modules. The blame
assignment says that the value witnessing the impedance mismatch may have
crossed the boundaries between elements in the set, and each crossing
checked the value's type in a shallow manner.

This ambiguity in Transient blame raises the question of how the rational
programmer should react when the language produces a blame set. Our answer
is that the rational programmer has at least two reasonable options. The first
one is to select the module that is added to the blame sequence first and
assign types to only that one---after all, if fully checked, the types of this
first module should be able to detect an impedance mismatch earlier in the
evaluation of a program than the later ones. The second option is to select the
module that is added to the blame sequence last, effectively interpreting the
blame sequence as a boundary-aware stack.

@@

\begin{quote}
\it A \emph{Transient-first blame trail}/\emph{Transient-last blame trail} is a sequence of scenarios
$\conf_0,...\conf_n$ of $\system$ where for all $0 \leq i \leq n - 1$,
$\conf_i \subset \conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i =
\first{\mblame{\system, \conf_i}}$ where $\first{\mblame{\system,
\conf}}$/$\last{\mblame{\system, \conf}}$ is the
first/last module, respectively,  that Transient adds to the blame sequence for $\conf$.
\end{quote}

% \begin{quote}
% \it A  is a sequence of scenarios
% $\conf_0,...\conf_n$ of $\system$ where for all $0 \leq i \leq n - 1$,
% $\conf_i \subset \conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i =
%   \last{\mblame{\system, \conf_i}}$ where $\last{\mblame{\system, \conf}}$ is the
% last module Transient adds to the blame sequence for $\conf$.
% \end{quote}

The definition of \emph{Transient exception trails} is analogous to that for
Natural. It is used to as a baseline for Transient-first and Transient-last
blame.

### 2021-03-04 writing/rational.tex
uniform editorial voice, mistakes, typos, size of text in figures, captions

Section~\ref{sec:why-rational} explains, and illustrates with an example,
how a migratory type setting helps with finding the source of an impedance
mismatch. Roughly speaking, it encourages the rational programmer to
equip a module with types if it is blamed in an error message. A sequence
of such steps make up a portion of the migratory path in the lattice of
type migration~\cite{tfgnvf-popl-2016}. The lattice, described in Section
~\ref{sub:stuff},  serves as the common substrate for the definition modes
of the rational programmer mentioned in section~\ref{sec:challenges}.

@@

Hence the research problem is how to make the modes of the rational programmer
comparable even when they correspond to different semantics and offer different kinds of information. Sections~\ref{sub:natural}
through~\ref{sub:erasure} describe the various modes of the rational programmer and explain
the extent to which they are comparable. Section~\ref{sub:experiment}
enhances their comparability with the programmer-effort dimension. Finally, the last
subsection states the experimental questions and gives an overview of the
experimental process.

>>>>

Section~\ref{sec:why-rational} explains how a migratory type setting helps with
finding the source of an impedance mismatch. Roughly speaking, it encourages the
rational programmer to equip a module with types if it is blamed in an error
message. A sequence of such steps make up a portion of the migratory path in the
lattice of type migration~\cite{tfgnvf-popl-2016}. The lattice serves as the
common substrate for the definition modes of the rational programmer (see
sec.~\ref{sub:stuff}).

@@

Hence the research problem is develop modes for the rational programmer and to make
them comparable even when they correspond to different semantics and offer
different kinds of information (see secs.~\ref{sub:natural}
through~\ref{sub:erasure}). Programmer effort relative to a fixed debugging
scenario introduces another dimension along which the modes become may be
compared (see sec.~\ref{sub:effort}). With these notions in place, it
becomes possible to state the experimental questions and gives an overview of
the experimental process (see sec.~\ref{sub:experiment}).

### 2021-03-04 writing/results.tex
uniform editorial voice, mistakes, typos, size of text in figures, captions

A machine with two Intel Xeon Gold 6258R processors (28 doubly-threaded cores each)
and 500GB of memory ran the experimental process from section~\ref{sec:rational} on the
sampled debugging scenarios from section~\ref{sec:sample}.  Each debugging scenario had a 4 minute timeout and a
memory limit of 6GB. The complete experiment required over 30,000 compute
hours.

\begin{wrapfigure}{r}{.40\textwidth}
  \includegraphics[width=0.40\textwidth]{./plots/success-bars}

  \vspace{1em}
  \begin{minipage}{0.4\textwidth}
  The estimated percentage of scenarios for which each mode succeeds in locating the bug.
  The upper bound margin of error is 0.02\%.
  \end{minipage}

  \caption{Success rates.}
  \label{fig:success-bars}
\end{wrapfigure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{./plots/avo-bars}

  \vspace{1em}
  \begin{minipage}{0.95\textwidth}
  Each plot depicts a head-to-head comparison of the entitled mode against every other mode.
  The portion above 0 (in green) is the estimated percentage of scenarios where the entitled mode is more useful
    than the other mode.
  The portion below 0 (in red) is the estimated percentage of scenarios
    where the entitled mode is less useful than the other mode.
  The upper bound margin of error is 0.02\%.
  \end{minipage}

  \caption{Usefulness comparisons.}
  \label{fig:avo-bars}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.32\textwidth]{./plots/TR-TR-stack-first-venn}
  \hfill
  \includegraphics[width=0.32\textwidth]{./plots/transient-newest-transient-stack-first-venn}
  \hfill
  \includegraphics[width=0.32\textwidth]{./plots/transient-oldest-transient-stack-first-venn}

  \vspace{1em}
  \begin{minipage}{0.95\textwidth}
  Each diagram shows the overlap of the successful scenarios for three modes.  
     For example, in the leftmost diagram, all three modes succeed on the
    same scenario 57.3\% of the time, only Natural and Natural exceptions
    succeed on 29.1\% of the scenarios, only Natural and Erasure succeed
    on 2.1\%, and Natural alone succeeds on 9\%.  The upper bound margin
    of error is 0.02\%.
  \end{minipage}

  \caption{Blame usefulness analysis.}
  \label{fig:success-venns}
\end{figure}

Figure~\ref{fig:success-bars} summarizes the overall success rates of
every mode.\footnote{The random mode behaves the same for
all three semantics, so this analysis considers it a single mode.}
The success rates illustrate a few points that underlie the
rest of the analysis.  The first notable piece of information from this
figure is that every mode has failed debugging scenarios, not just
Erasure. This should not come as a surprise to the astute reader.  Running
a rational programmer mode on a scenario may result in an exception
that carries no useful information about which module to equip with types next. For instance the stack trace of the exception may not
contain frames from any untyped module of the program. This can happen at any point
along a blame trail, causing it to fail.

@@

In Transient, similar to Natural, most failures are due to unhelpful
exception information (1851 for both Transient first and last blame).
However, Transient also has a substantial number of failures because
scenarios hit the time and memory limits of the experiment
(\textasciitilde770 scenarios).  Additionally, there are nearly 1,000
cases where Transient reports an empty blame set, leaving the rational
programmer without hints about how to proceed.
Sections~\ref{sec:threat:transient} and ~\ref{sec:threat:transient2}
explain both of these causes of failure for Transient in detail and how
they affect the
experiment.

@@

Figure~\ref{fig:avo-bars} depicts a head-to-head comparison of every
mode's performance against every other mode (except Random). The
comparison  answers the four questions from section~\ref{sub:experiment}. 
Each plot shows the proportion of scenarios where one mode performs
better or worse than each other mode.  In particular, each bar above zero
represents the proportion where the plot's entitled mode succeeds and the
mode on the x-axis fails; the corresponding bar below zero represents the
proportion of the inverse case.  For example, the plot titled ``Natural''
shows that Natural outperforms Natural exceptions in about 12\% of the
scenarios, and the inverse (Natural performs worse than Natural
exceptions) never happens.  Similarly, the plot titled ``Transient last
blame'' shows that Transient last blame outperforms Natural exceptions
 in about 12\% of the scenarios, but conversely it performs worse
than Natural exceptions in about 2\% of the scenarios.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{./plots/bt-lengths-table}

@@

  \caption{Programmer effort.}
  \label{fig:effort-table}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{./plots/bt-length-comparisons}

  \vspace{1em}
  \begin{minipage}{0.95\textwidth}
  Each plot depicts the distribution of scenarios with trail length
    differences ranging from -3 to 3.  A -x difference denotes that
    the first mode's trail is x steps shorter than the second mode's
    trail for the same scenario, and a positive difference denotes the
    inverse.  Only scenarios where both modes succeed are considered.  The
    10|85 on the y-axis indicates that the axis is truncated between 10
    and 85\%.
  \end{minipage}

  \caption{Effort comparisons.}
  \label{fig:effort-comparisons}
\end{figure}

In order to better understand the answers the experiment provides for
questions $Q_1$ to $Q_3$, figure~\ref{fig:success-venns} depicts an
analysis of the success of each semantics in comparison to Erasure.
Specifically, the figure shows one Venn diagram per mode of the rational
programmer that uses blame.  Each diagram shows the overlap of successful
scenarios for  the blame mode, its corresponding exception mode, and Erasure.  For
example, the leftmost diagram (Natural) shows that all three modes
succeed on 59.5\% of the  scenarios, only Natural and Natural exceptions
succeed on 25.7\% of the scenarios, only Natural and Erasure succeed on
2.1\%, and Natural alone succeeds on 9.7\%.  This analysis highlights
the success trade-offs each semantics offers against Erasure, with and
without blame. For instance, the analysis
for Natural clearly illustrates that, when choosing between Natural blame,
Natural exceptions, and Erasure, Natural blame is the absolutely most
successful: all of the successes of the other two modes are subsets of
Natural's successful scenarios.  On the other hand, Transient's blame
modes fare similarly but the choice is not so clear-cut.

@@

A second point of interest is that the exceptions of the Erasure semantics
either help the rational programmer immediately or the rational programmer
fails to debug a scenario altogether (in red).  This is expected; adding
type annotations to an Erasure program doesn't change it's behavior, so if
the type checker doesn't reject the program with the new annotations, the
rational programmer is stuck.  In other words, an exception from the
runtime has to point to the buggy module with the first try. Otherwise the
rational programmer types an irrelevant module, runs the program again,
and the same exception points again to the already typed module.

Figure~\ref{fig:effort-comparisons} provides head-to-head comparisons of
effort.  The comparison between two modes boils down to the difference in
length between their trails for all scenarios where they both succeed.
Hence, each plot in the figure shows the distribution of scenarios with
length differences ranging from -3 (the first mode's trail is 3 steps
shorter than the second's) to 3 (the first mode's trail is 3 steps longer
than the second's).  The figure offers several insights about how  modes
compare in terms of effort that complement the insights about how they
compare in terms of success rates from figure~\ref{fig:avo-bars}.  First,
Natural blame never produces shorter trails than Natural exceptions, and
sometimes (albeit rarely) produces slightly longer ones.  Hence, the
experiment provides evidence that blame helps the rational programmer
debug more scenarios but does not shorten the debugging process compared
to exceptions. A second insight is that Natural relatively often (around 10\% of the
scenarios) produces shorter trails than both Transient blame modes, and
sometimes the trails are significantly shorter.  Third, Transient last
blame often produces shorter trails than Transient first blame, which is a
particularly interesting distinction in light of their nearly
indistinguishable success rates.  Finally, Transient's blame modes also
share the characteristic with Natural that blame sometimes produces longer
trails than their corresponding exception modes.

>>>>

\begin{wrapfigure}{r}{.40\textwidth} \footnotesize 
  \vspace{-1.5em}
  \includegraphics[width=0.38\textwidth]{./plots/success-bars}
  \vspace{1em}
  \begin{minipage}{0.35\textwidth} \raggedright
%   The estimated percentage of scenarios for which each mode succeeds in locating the bug.
   The upper bound margin of error is 0.02\%.
  \end{minipage}
\vspace{-2em}
  \caption{Percentage rates of success} \label{fig:success-bars}
\end{wrapfigure}

The test bed for executing the experimental process utilizes a machine with two
Intel Xeon Gold 6258R processors (28 doubly-threaded cores each) and 500GB of
memory.  Each debugging scenario had a 4 minute timeout and a 6GB memory
limit. Running the experiment on all debugging scenarios took over
30,000 compute hours.

\begin{figure} \footnotesize 
  \centering
  \includegraphics[width=\textwidth]{./plots/avo-bars}

  \vspace{1em}
  \begin{minipage}{0.95\textwidth}
      Each plot depicts a head-to-head comparison of the mode named above the
      plot vs. every other mode.  The (green) portion above 0 is the estimated
      percentage of scenarios where the named mode is more useful than the
      other.  The (red) portion below 0 is the estimated percentage of scenarios
      where the named mode is less useful than the other.  The upper bound
      margin of error is 0.02\%.
  \end{minipage}

  \caption{Usefulness comparisons}
  \label{fig:avo-bars}
\end{figure}

\begin{figure}\footnotesize 
  \centering
  \includegraphics[width=0.32\textwidth]{./plots/TR-TR-stack-first-venn}
  \hfill
  \includegraphics[width=0.32\textwidth]{./plots/transient-newest-transient-stack-first-venn}
  \hfill
  \includegraphics[width=0.32\textwidth]{./plots/transient-oldest-transient-stack-first-venn}

  \vspace{1em}
  \begin{minipage}{0.95\textwidth}

     Each diagram shows the overlap of the successful scenarios for three modes.
     For example, in the leftmost diagram, all three modes succeed on the same
     scenario 57.3\% of the time, only Natural and Natural exceptions succeed on
     29.1\% of the scenarios, only Natural and Erasure succeed on 2.1\%, and
     Natural alone succeeds on 9\%.  \\ The upper bound margin of error is 0.02\%.

  \end{minipage}

  \caption{Blame usefulness analysis}
  \label{fig:success-venns}
\end{figure}

% \footnote{The random mode behaves the same for all three semantics, so this analysis considers it a single mode.}

Figure~\ref{fig:success-bars} summarizes the overall success rates of
every mode.
The success rates illustrate a few points that underlie the
rest of the analysis.  The first notable piece of information from this
figure is that every mode has failed debugging scenarios, not just
Erasure. This should not come as a surprise to the astute reader.  Running
a rational programmer mode on a scenario may result in an exception
that carries no useful information about which module to equip with types next. For instance the stack trace of the exception may not
contain frames from any untyped module of the program. This can happen at any point
along a blame trail, causing it to fail.

@@

In Transient, similar to Natural, most failures are due to unhelpful exception
information (1851 for both Transient first and last blame).  However, Transient
also has a substantial number of failures because scenarios hit the time and/or
memory limits of the experiment (\textasciitilde770 scenarios).  Additionally,
there are nearly 1,000 cases where Transient reports an empty blame set, leaving
the rational programmer without hints about how to proceed.
Sections~\ref{sec:threat:transient} and ~\ref{sec:threat:transient2} address
these causes of failure for Transient and how they affect the experiment.

@@

Figure~\ref{fig:avo-bars} depicts a head-to-head comparison of every
mode's performance against every other mode (except Random). The
comparison  answers the four questions from section~\ref{sub:experiment}. 
Each plot shows the proportion of scenarios where one mode performs
better or worse than each other mode.  In particular, each bar above zero
represents the proportion where the plot's named mode succeeds and the
mode on the x-axis fails; the corresponding bar below zero represents the
proportion of the inverse case.  For example, the plot titled ``Natural''
shows that Natural outperforms Natural exceptions in about 12\% of the
scenarios, and the inverse (Natural performs worse than Natural
exceptions) never happens.  Similarly, the plot titled ``Transient last
blame'' shows that Transient last blame outperforms Natural exceptions
 in about 12\% of the scenarios, but conversely it performs worse
than Natural exceptions in about 2\% of the scenarios.

@@

\begin{figure} \footnotesize \centering
  \includegraphics[width=\textwidth]{./plots/bt-lengths-table}

@@

  \caption{Programmer effort} \label{fig:effort-table}
\end{figure}

\begin{figure}\footnotesize \centering
  \includegraphics[width=\textwidth]{./plots/bt-length-comparisons}

  \vspace{1em}
  \begin{minipage}{0.95\textwidth}

    Each plot depicts the distribution of scenarios with trail length
    differences ranging from -3 to 3.  A -x difference denotes that the first
    mode's trail is x steps {\em shorter\/} than the second mode's trail for the
    same scenario; a positive difference denotes the inverse.  Only
    scenarios where both modes succeed are considered.  The 10|85 on the y-axis
    indicates that the axis is truncated between 10 and 85\%.

  \end{minipage}

  \caption{Effort comparisons} \label{fig:effort-comparisons}
\end{figure}

An alternative way to understand the answers for questions $Q_1$ to $Q_3$, is to
analyze the success of each semantics in comparison to
Erasure. Figure~\ref{fig:success-venns} depicts the results of this analysis.
Specifically, the figure shows one Venn diagram per mode of the rational
programmer that uses blame.  Each diagram shows the overlap of successful
scenarios for the blame mode, its corresponding exception mode, and Erasure.
For example, the leftmost diagram (Natural) shows that all three modes succeed
on 59.5\% of the scenarios, only Natural and Natural exceptions succeed on
25.7\% of the scenarios, only Natural and Erasure succeed on 2.1\%, and Natural
alone succeeds on 9.7\%.  This analysis highlights the success trade-offs each
semantics offers against Erasure, with and without blame. For instance, the
analysis for Natural clearly illustrates that, when choosing between Natural
blame, Natural exceptions, and Erasure, Natural blame is the absolutely most
successful: all of the successes of the other two modes are subsets of Natural's
successful scenarios.  On the other hand, Transient's blame modes fare similarly
but the choice is not so clear-cut.

@@

Second, the exceptions of the Erasure semantics either help the rational
programmer immediately or the rational programmer fails to debug a scenario
altogether (in red).  This is expected; adding type annotations to an Erasure
program does not change its behavior, so if the type checker does not reject the
program with the new annotations, the rational programmer is stuck.  In other
words, an exception from the runtime has to point to the buggy module with the
first try. Otherwise the rational programmer types an irrelevant module, runs
the program again, and the same exception points again to the already typed
module.

Figure~\ref{fig:effort-comparisons} provides head-to-head comparisons of
effort.  The comparison between two modes boils down to the difference in
length between their trails for all scenarios where they both succeed.
Hence, each plot in the figure shows the distribution of scenarios with
length differences ranging from -3 (the first mode's trail is 3 steps
shorter than the second's) to 3 (the first mode's trail is 3 steps longer
than the second's).  The figure offers several insights about how  modes
compare in terms of effort that complement the insights about how they
compare in terms of success rates from figure~\ref{fig:avo-bars}.  First,
Natural blame never produces shorter trails than Natural exceptions, and
sometimes (albeit rarely) produces slightly longer ones.  Hence, the
experiment provides evidence that blame helps the rational programmer
debug more scenarios but does not shorten the debugging process compared
to exceptions. Second, Natural relatively often (around 10\% of the
scenarios) produces shorter trails than both Transient blame modes, and
sometimes the trails are significantly shorter.  Third, Transient last
blame often produces shorter trails than Transient first blame, which is a
particularly interesting distinction in light of their nearly
indistinguishable success rates.  Finally, Transient's blame modes also
share the characteristic with Natural that blame sometimes produces longer
trails than their corresponding exception modes.

### 2021-03-04 writing/sample.tex
touch ups in section 7
uniform editorial voice, mistakes, typos, size of text in figures, captions

As is, the chosen mutators generate a huge number of debugging scenarios:
approximately 1 million for the chosen benchmarks. This number of scenarios is
far too large to even identify the interesting ones among them.  Furthermore
this population is heterogeneous; scenarios come from different
mutants, and the mutants are the result of different mutators applied to each
benchmark.

To become computationally feasible and statistically sound, the experiment employs
stratified uniform random sampling with three levels of stratification.
The first two levels group mutants first by benchmark and then by
mutator, making sure that the sample within each benchmark reflects the diversity of
mutants with respect to the mutators that generated them.  Specifically, the
experiment samples 80 interesting mutants per benchmark,
evenly-distributed across all of the mutators that contribute mutants for the
benchmark. Some benchmarks have less than  80 mutants with interesting scenarios,
in which case the only choice is to take them all.
The result is a total of 756 interesting mutants across all benchmarks.
Finally, the third level of sampling draws 96 random interesting debugging
scenarios from each configuration lattice with replacement. The
final sample thus consists of 72,192 interesting scenarios.

>>>>

As is, the chosen mutators generate approximately one million debugging
scenarios for the chosen benchmarks. This number of scenarios is far too large
to even identify the interesting ones among them.  Furthermore this population
is heterogeneous; scenarios come from different mutants, and the mutants are the
result of different mutators applied to each benchmark.

To become computationally feasible and statistically sound, the experiment employs
stratified uniform random sampling with three levels of stratification.
The first two levels group mutants first by benchmark and then by
mutator, making sure that the sample within each benchmark reflects the diversity of
mutants with respect to the mutators that generated them.  Specifically, the
experiment samples 80 interesting mutants per benchmark,
evenly-distributed across all of the mutators that contribute mutants for the
benchmark. Some benchmarks have less than  80 mutants with interesting scenarios,
in which case the only choice is to include them all.
The result is a total of 756 interesting mutants across all benchmarks.
Finally, the third level of sampling draws 96 random interesting debugging
scenarios from each configuration lattice with replacement. The
final sample thus consists of 72,192 interesting scenarios.

%% MF: I don't get this number ^^^^
% > (* 756 96)
% 72576

### 2021-03-04 writing/why-rational.tex
uniform editorial voice, mistakes, typos, size of text in figures, captions

The general challenge of evaluating blame is a methodological one. Unlike most
current research on programming languages, the question seems to call for
empirical research similar to that of human-computer interaction. At the same
time, a significant evaluation result demands a large amount of
data. As~\citet{lksfd-popl-2020} recently demonstrated, the way around this
dilemma is to simulate a programmer's activities algorithmically on a large set
of programming scenarios.

@@

The idea is best illustrated with an example in Typed Racket's migratory type
system. Imagine a code base with dozens of modules in plain Racket. A developer
who opens a module for maintenance purposes must study the module's design and,
as part of the process, is bound to re-construct the types that went into the
module's creation.  To help future maintainers, the developer should report
these insights as type annotations. Over time, the code base migrates into a mix
of typed and untyped modules. As~\citet{tfffgksst-snapl-2017} reported, though,
it is at least equally common that developers add typed modules that depend on
the existing modules in the code base.

@@

Assuming the rational programmer trusts the type of {\tt na-client}, the next
step is to inspect the {\tt layer} module and to equip it with type
annotations. The result is the \typecolor\ module in the middle, and {\tt
main}'s import is now re-directed there by {\tt (submod ".." layer-typed)}. Running the
modified program yields a different error message:
\begin{verbatim}
    neg-abs: broke its own contract
      promised: (<=/c 0)
      produced: 10
      in: (-> any/c (<=/c 0))
      contract from: (interface for neg-abs)
      blaming: (interface for neg-abs)
       (assuming the contract is correct)
\end{verbatim}
The last step of the rational programmer is to assign types to the {\tt server}
module and to re-direct the import of {\tt layer-typed} to {\tt (submod ".."
server-typed)}. At this point the type checker objects to the conjectured type
of {\tt neg-abs}---and the source of the impedance mismatch is found. How to fix
it is a separate question.

>>>>

The general challenge of evaluating blame is a methodological one. Unlike most
current research on programming languages, the question seems to call for
empirical studies similar to those of human-computer interaction. At the same
time, a significant evaluation result demands a large amount of
data. As~\citet{lksfd-popl-2020} recently demonstrated, the way around this
dilemma is to simulate a programmer's activities algorithmically on a large set
of programming scenarios.

@@

The idea is best illustrated with an example in Typed Racket's migratory type
system. Imagine a code base with dozens of modules in plain Racket. A developer
who opens a module for maintenance purposes must study the module's design and,
as part of the process, is bound to re-construct the types that went into the
module's creation.  To help future maintainers, the developer should report
these insights as type annotations. Over time, the code base migrates into a mix
of typed and untyped modules. As~\citet{tfffgksst-snapl-2017} report though,
it is equally common that developers add typed modules that depend on
the existing modules in the code base.

@@

Assuming the rational programmer trusts the type of {\tt na-client}, the next
step is to inspect the {\tt layer} module and to equip it with type
annotations. The result is the \typecolor\ module in the middle, and {\tt
main}'s import is now re-directed there by {\tt (submod ".." layer-typed)}. As
predicted by the theory, running the modified program yields a different error message:
\begin{verbatim}
    neg-abs: broke its own contract
      promised: (<=/c 0)
      produced: 10
      in: (-> any/c (<=/c 0))
      contract from: (interface for neg-abs)
      blaming: (interface for neg-abs)
       (assuming the contract is correct)
\end{verbatim}
The last step of the rational programmer is to assign types to the {\tt server}
module and to re-direct the import of {\tt layer-typed} to {\tt (submod ".."
server-typed)}. At this point the type checker objects to the conjectured type
of {\tt neg-abs}---and the source of the impedance mismatch is found. How to fix
it is a separate question.
